---
title: "Homework 4 Final Draft"
format: html
editor: visual
---

## Set Up

```{r}
#install packages
install.packages("lterdatasampler")

# should haves
library(tidyverse)
library(here)
library(lterdatasampler)

# would be nice to have
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
```

# Problem: How does fish length predict fish weight for trout perch (across all sample years)?

### #1: Null and Alternative Hypotheses

Null Hypothesis: There is no significant relationship between fish length and fish weight in the trout perch samples.

Alternative Hypothesis: There is relationship between fish length and fish weight in the trout perch samples, meaning that fish weight can reasonably be predicted using the fish length.

### #2: Visualization of Missing Data

Read in data and filter so that only Trout Perch data are included. Then, visualize the missing data using the naniar package.

```{r}
data <- read.csv("data/knb-lter-ntl.6.34/ntl6_v12.csv") #read in data

trout_perch_data <- data %>% #Create new item for the fitered data including only Trout Perch observations
  filter(spname == "TROUTPERCH")  #Filter so that only observations with species name "Trout Perch" are included

library(naniar) 
vis_miss(trout_perch_data) #Visualize missing data in the filtered data set

```

The missing data is very relevant to the hypotheses, because we are evaluating the extent to which fish length can be predicted to fish weight, and the weight variable is missing 41% of its data.

### #3: Visualizing Data and Establishing Potential for Relationship

Make a scatter plot for reference:

```{r}
ggplot(data = trout_perch_data, aes(x = length, y = weight))+ #create a plot based on the trout_perch_data dataset, with the x variable being the fish lenght, and the y variable being the fish weight
  geom_point() #Establishes the type of plot as a scatter plot
```

The scatter plot seems to indicate a relatively strong relationship between fish length and fish weight. Further analysis will be needed to check that this assumption holds true.

### #4: Checking Assumptions of Test with Diagnostic Plots

```{r}
modeltrout <- lm(weight ~ length, data = trout_perch_data) 
#creates a linear regression model using the Trout Perch Dataset. Establishes the potential dependent variable as weight and the potential intependent variable as length

par(mfrow = c(2, 2)) #Organizes the plots in a 2x2 grid for easier visualization
plot(modeltrout) #Plots the linear regression model that was just created
```

### #5: Analysis of Diagnostic Plots:

-   Residuals vs. Fitted:

-   Q-Q Residuals:

-   Scale-Location:

-   Residuals vs. Leverage:

### #6 - #7: Results of Summary(), ANOVA Table, and Summarized ANOVA Table

```{r}
model_summary <- summary(modeltrout) #Stores summary as an object
model_squares <- anova(modeltrout) #Stores ANOVA table as an object

model_summary
```

```{r}
model_squares
```

```{r}
#Summarize the ANOVA table 

model_squares_table <- tidy(model_squares) %>% 
  # round sum of squares and mean squares columns
  mutate(across(sumsq:meansq, ~ round(.x, digits = 3))) %>% 
  # round the F-statistic to 1 digit
  mutate(statistic = round(statistic, digits = 1)) %>% 
  # replace the p value with < 0.001 so as to not appear in scientific notation
  mutate(p.value = case_when(
    p.value < 0.001 ~ "< 0.001"
  )) %>% 
  # rename the length (independent) variable
  mutate(term = case_when(
    term == "length" ~ "Fish Length",
    TRUE ~ term
  )) %>% 
  # make data frame a flextable object
  flextable() %>% 
  # change the header labels to be more explanatory
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")

model_squares_table
```

### #8: Explaining ANOVA Table Results

The ANOVA table explains

### #9: Summary of Results

### #10: Finalized Model Predictions with Confidence Intervals and Underlying Scatterplot Data

Visualizing the results

```{r}
predictions <- ggpredict(modeltrout, terms = "length") 
#uses ggpredict to establish model predictions
#assigns to term "predictions"

predictions
```

```{r}
plot_predictions <- ggplot(data = trout_perch_data, #plot the data and predictions.
                           aes(x = length, y = weight)) + #x axis represents fish length, y axis repesents fish weight
 
  
   #plot the data as a scatterplot
  geom_point() +
 
   # plot predictions as a line
  geom_line(data = predictions, 
            aes(x = x, y = predicted), #establishes x axis as x, and y axis as predicted.
            color = "red", linewidth = 1) + #assigns the color red to the prediction line
  
  
  #plot the 95% confidence interval from ggpredict
  geom_ribbon(data = predictions, 
              aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  
  
  # theme and clearer labels
  theme_bw() +
  labs(x = "Trout Perch Length",
       y = "Trout Perch Weight")

plot_predictions
```

The above figure displays the observed Trout Perch weight and length observations (black dots) in a scatter plot, as well as the predicted values (red line) and the 95% confidence interval (light grey). The majority of the observed data seems to align near the predicted values. The narrow confidence intervals also indicate that the prediction model does indeed fit the data well. See answer to "#9: Summary of Results" for more in depth explanation and contextualization of this visualization and its results.
